---
title: "The prolific.api package"
author: "Simon Lenau"
date: "`r Sys.Date()`"
output:
rmarkdown::html_vignette:
    toc:yes
    code_folding:hide
vignette: >
  %\VignetteIndexEntry{The prolific.api package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
params:
  EVAL: !r identical(Sys.getenv("NOT_CRAN"), "true")
---

```{r, include = FALSE}
stopifnot(require(knitr))
options(width = 90)
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = !TRUE,
  comment = "#>",
  dev = "jpeg",
  dpi = 100,
  fig.asp = 0.8,
  fig.width = 5,
  out.width = "60%",
  fig.align = "center"
)
library(prolific.api)
```

The **prolific.api** package provides functionalities for creating and managing empirical crowd-sourcing studies on the 
[prolific.co](https://www.prolific.co/) platform by using the platform's API access.

## Introduction


***Note:***
An
[API token](https://docs.prolific.co/docs/api-docs/public/#section/Authentication) 
is required for accessing the API of 
[prolific.co](https://www.prolific.co/). 
For more details, see the section on *[obtaining an API token](#get-token)*.



## Core Functionalities

The main functionalities in the **prolific.api** package are provided in three
`ReferenceClasses`:

| Class                  | Functionality                                         |
| :--------------------- | :---------------------------------------------------- |
| `api_access`           | - Access the API<br>- Submit and retrieve information |
| `prolific_study`       | - Set up and modify studies                           |
| `prolific_prescreener` | - Define the group of eligible participants           |

*In general, all fields and methods of these classes are available in a RefClass as well as S4 object style (see examples).*
The core functionalities are summarized below.

### Submitting and retrieving information: The `api_access` class

The `api_access` class is designed for interacting with 
[Prolific's API](https://docs.prolific.co/docs/api-docs/public/).
The central method to achieve this is `access`, which can be used
to exchange (retrieve and submit) information with the API.

#### Setting up `api_access`

An `api_access` object can be created by 

```{r include = FALSE,eval=TRUE}
prolific_api_access <- api_access(api_token = readLines("~/.prolific_api/R_package_prolific.api_token", warn = FALSE))
```

```{r eval = FALSE}
prolific_api_access <- api_access(api_token = "<api_token>")
```

The `api_token` is the only information that **must** be specified, for the API acces to work. 
All other settings rarely require adjustment.
To check whether the token is valid,  use the `check_authorization` method:
```{r eval = TRUE}
check_authorization(prolific_api_access)
```

The section on 
*[obtaining an API token](#get-token)* 
describes how to obtain a valid token.

#### Using `api_access`

The API access is then carried out using the `access` method.
An example for retrieving data from [prolific.co](https://www.prolific.co/) is

```{r eval = FALSE}
access(prolific_api_access,
  method = "get",
  endpoint = "users/me"
)
```
which shows some information about the account you are accessing the API with.
A simple example where data is actually sent to the API is

```{r eval = FALSE}
access(prolific_api_access,
  endpoint = "study-cost-calculator",
  method = "post",
  data = list(
    reward = 100,
    total_available_places = 5
  )
)
```

which calculates the cost  (including fees and taxes)  of a study where `5` participants are paid `1` € each.


The above examples use 
`
  method = "get"
`
and
`
method = "post"
`, respectively.
The list of available options for the `method` argument in `access()` is:

| Method     | Function                                   |
| :--------- | :----------------------------------------- |
| `"get"`    | Retrieving endpoint / data                 |
| `"post"`   | Create endpoint / send data                |
| `"patch"`  | Change endpoint using a delta (difference) |
| `"put"`    | Replace endpoint                           |
| `"delete"` | Delete endpoint                            |

More realistic examples are provided in the [examples](#examples) section below.

A list of further endpoints is provided in the 
[vignette on Prolific's API endpoints](prolificapi-endpoint.html)
as well as 
[Prolific's API documentation](https://docs.prolific.co/docs/api-docs/public) itself.

### Set-up and change studies: The `prolific_study` class

The `prolific_study` class provides a lightweigt interface for
creating, and managing studies on [prolific.co](https://www.prolific.co/) using `R`.
There are a lot of options to be chosen from when setting up such a study, but let's start with a simple example.

#### Creating a `prolific_study`

A minimal specification for creating a `prolific_study` contains the following information:

```{r eval = TRUE}
new_study <- prolific_study(
  name = "<Publicly visible study name>",
  description = "<Publicly visible study description>",
  estimated_completion_time = 1,
  reward = 1,
  external_study_url = "https://www.link_to_my_study.com",
  completion_code = "123",
  total_available_places = 10
)
```


The information that is presented to the potential participants 
contains the study's 

- **name**,
- **description**,
- **estimated completion time** and 
- **reward**.

People deciding to take part in a study are then redirected to an

- **external study url**

where the study is conducted.
Once they completed your study, you should provide them with a

- **completion code** 

and send them back to [prolific.co](https://www.prolific.co/).
Participant compensation is then based on the completion code by checking
whether a participant obtained the correct completion code
after completing the study.
^[Participant compensation does not happen automatically, so you still can check which participants are compensated, e.g. in case of erroneous completion codes.]
The  (maximum) the number of people that can access your study via [prolific.co](https://www.prolific.co/) is determined by the

- **total available places**.


#### Checking the Validity of a `prolific_study`

``` {r eval = TRUE}
validity_check(new_study)
```

#### Updating Fields of a `prolific_study`


#### Submitting, Updating and Deleting a `prolific_study`

A study that is created or changed in `R` needs to be submitted to [prolific.co](https://www.prolific.co/) 
once

``` {r eval = FALSE}
prolific_api_access$access(
  endpoint = "studies",
  method = "post",
  data = new_study
)
```

``` {r eval = TRUE, include = FALSE}
prolific_api_access$access(
  endpoint = c("studies", new_study$id),
  method = "delete"
)
```

### Defining eligible participants: The `prolific_prescreener` class 

A list of all currently available prescreeners is provided in the
[Prolific prescreeners vignette](prolificapi-prescreeners.html)


## Examples {#examples}

```{r setup}
library(prolific.api)
```



## Obtaining an API token {#get-token}

The **prolific.api** package requires 
an 
[API token](https://docs.prolific.co/docs/api-docs/public/#section/Authentication) 
for accessing the API of 
[prolific.co](https://www.prolific.co/).


To get such a token, you first need a researcher account on 
[prolific.co](https://www.prolific.co/).
When logging in to this account, you can obtain the *workspace-specific* token
by selecting the workspace you would like to work in, for which 
the tokens can be managed in the `Settings -> Go to API token page` menu.


